---
title: "No Such Thing as a 4-2-3-1"
author: "Aditya Kothari"
date: "May 2019"
output: html_document
---

```{r Setup, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'asis'}
   
rm(list = ls())

library(data.table)
library(emdist)
library(snow)
library(snowfall)
library(ggplot2)
library(scales)
library(knitr)
library(clue)
library(dendextend)
theme_set(theme_bw(12))

```

```{r ParametersFunctions, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'asis'}

cPostCodeFolderLocation = '/media/ask/Data/Personal/Projects/Personal/AnalysingFootballData/EMDRelated/'
cPostDataFolderLocation = '/media/ask/Data/Personal/Projects/Personal/NoSuchThingAs4231'

source(paste0(cPostCodeFolderLocation, '/Common.R'))
read_chunk(paste0(cPostCodeFolderLocation, '/CommonChunksClusteringFormations.R'))

iChunks = 100

```


```{r LoadingData, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'asis'}

load(paste0(
   cPostDataFolderLocation,
   '/Data/FormationMatchingData/ProcessedData.Rdata'
))

load(paste0(
   cPostDataFolderLocation,
   '/Data/FormationMatchingData/ClusteredData.Rdata'
))

dtDistancMatchFileNames = merge(
   dtDistancMatchFileNames,
   dtPasses[, list(SNO1 = SNO[1]), list(matchId1 = as.character(matchId), teamId1 = as.character(teamId))],
   c('matchId1','teamId1')
)

dtDistancMatchFileNames = merge(
   dtDistancMatchFileNames,
   dtPasses[, list(SNO2 = SNO[1]), list(matchId2 = as.character(matchId), teamId2 = as.character(teamId))],
   c('matchId2','teamId2')
)

```
# {.tabset .tabset-fade .tabset-pills}

This is a very very WIP post. It has only EPL comparisons right now and I'm adding data from other leagues to get more generic results.

## Motivation

Formations and player roles are a very fuzzy concept. We simplify things into 4-2-3-1s and 4-3-3s and right backs and centre forwards but within those simplifications there are many nuances to how different teams and different players function.

In this post, I try to quantify these and find teams who are set up similarly in various matches.

## Methodology

Refer to <a href="https://thecomeonman.github.io/SpatialSimilaritiesBetweenPlayers/">my post on similar players based on their spatial distributions</a>.

## Clustering team-matches

The teams are clustered based on the max distance between players.

```{r Clustering, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'asis'}


```

Each node at the end of this graph is a particular team playing in a particular match.

```{r ClusteringResults, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 5, results = 'asis'}


```

Cluster descriptions:

* Cluster 1 is 3-4-2-1

* Cluster 2 is 3-4-2-1 with a mix of various other 3 at the back formations

* Cluster 3 is 4-3-3

* Cluster 4 to 7 are a mix of various 4 at the back formations with a strong 4-2-3-1 element in all of them.

* Cluster 8 seems to not have an underlying link with the formation.


## Cluster Examples {.tabset .tabset-fade .tabset-pills}

I've picked a match from each of the clusters, and overlaid that match with data from all the player pairings with that match in that cluster. That gives a rough idea of how the teams played in that cluster.

I've also added some comparisons between clusters which feel slightly similar.


```{r ClusterSetting1, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'asis'}

iClusterToDemonstrate = 1

```

### Cluster `r iClusterToDemonstrate`  {.tabset .tabset-fade .tabset-pills}

#### Cluster Example

```{r ClusterExploration1, ref.label = 'ClusterExplorationTemplate', cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 15, results = 'asis'}
```

#### Comparisons {.tabset .tabset-fade .tabset-pills}

##### Clusters 1 and 2

```{r ClusterComparison12Setting, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'asis'}

iClusterToCompare = 2
cFormationToCompare = '3421'

```

```{r ClusterComparison12, ref.label = 'ClusterComparisonTemplate', cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 15, results = 'asis'}
```


##### Clusters 1 and 3

```{r ClusterComparison13Setting, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'asis'}

iClusterToCompare = 3
cFormationToCompare = '3421'

```

```{r ClusterComparison13, ref.label = 'ClusterComparisonTemplate', cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 15, results = 'asis'}
```






```{r ClusterSetting2, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'asis'}

iClusterToDemonstrate = 2

```

### Cluster `r iClusterToDemonstrate`  {.tabset .tabset-fade .tabset-pills}

#### Cluster Example

```{r ClusterExploration2, ref.label = 'ClusterExplorationTemplate', cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 15, results = 'asis'}
```










## To Do

I plan to add data from La Liga, Bundesliga, Seria A, Ligue 1, and the Championship from 2017/18. A bigger set of matches with more diverse strategies might lead to different, and more general clusters?